---
title: "Week 2 solution"
author: "Kathrine Schultz-Nielsen"
date: "8/2/2023"
output:
---

```{r}
pacman::p_load(rstanarm,tidyverse, ggpubr, rootSolve)

```

# Warm-up: practicing vector and matrix notation
Consider these the following predictions
$$
\begin{aligned}
& \hat{y}_1=47=50+3.0 *(-1) \\
& \hat{y}_2=-50=50*(-1)+3.0 * 0 \\
& \hat{y}_3=109=50*2+3.0 * 3
\end{aligned}
$$
```{r}
50+3.0 *(-1)
(-1)*50+3.0 * 0
(2)*50+3.0 * 3
```
Try to write the prediction in vector notation. If you further want to challenge yourself try Matrix notation as well. (if you are stuck have a look at pages 36-37 in ROS). It might be helpful to do this exercise by hand.

## Solution
Vector notation
$$
\hat{y}=\left(\begin{array}{c}
47 \\
-50 \\
109
\end{array}\right)=\left(\begin{array}{c}
50+3.0 *(-1) \\
50*(-1)+3.0 * 0 \\
50*2+3.0 * 3
\end{array}\right)
$$
Matrix notation
$$
\hat{y}=\left(\begin{array}{c}
43.3 \\
46.3 \\
55.3
\end{array}\right)=\left(\begin{array}{cc}
1 & -1 \\
-1 & 0 \\
2 & 3
\end{array}\right)\left(\begin{array}{c}
50 \\
3.0
\end{array}\right)
$$

# 3.1
```{r}
# Computing the proportion of respondents in the sample who support higher taxes

# 'Manual' way
support = round((200*0.5 + 250*0.6 + 300*0.4 + 250*0.3)/(200 + 250 + 300 + 250), 4)
sprintf('Proportion supporting higher taxes: %s',support)

# Using R-dataframe & tidyverse
poll_df <- data.frame(Age = c("19:29","30:44","45:64","65+"), People = c(200,250,300,250), 
           percentage = c(0.5,0.6,0.4,0.3))
poll_df = summarise(poll_df, sum(People*percentage)/sum(People))

print(poll_df)
```

# 3.3
```{r}
# Create a sequence of numbers between -10 and 10 incrementing by 0.1.
x <- seq(-10, 10, by = .1)

# Choose the mean as 2.5 and standard deviation as 0.5.
y <- dnorm(x, mean = 0, sd = 1)
plot(x, y, type = "l", main = 'Varying the standard deviation', ylab = 'density')


for (sd_val in 2:4){
    x <- seq(-10, 10, by = .1)
    lines(x, dnorm(x, 0, sd_val))
}

y2 <- dnorm(x, mean = -2, sd = 1)
plot(x, y2, type = "l", main = 'Varying the mean', ylab = 'density')
for (mean_val in -1:2){
    x <- seq(-10, 10, by = .1)
    lines(x, dnorm(x, mean_val, 1))
  }
```


```{r}
# Using ggplot and simulation
#Simulate numbers for a density distribution
density_df <- data_frame(density_values = c(), mean_dens = c(), sd_dens = c())

# Simulating standard deviations 2,3 and 4
for (sd_val in 2:4){
  # looping over mean values from 10 to 15
  for (mean_val in 10:15){
    temp <- rnorm(1000, mean = mean_val,  sd = sd_val)
    mean = rep(mean_val,1000)
    sd = rep(sd_val, 1000)
    
    density_df <- rbind(density_df, data_frame(density_values = temp , mean_dens =  mean, sd_dens = sd))
  }
}
ggplot(density_df, aes(x = density_values, color = as.factor(mean_dens))) + geom_density(alpha = 0.4) + facet_grid(~as.factor(sd_dens)) + theme_classic()
```

# 3.4 
```{r}
# Poisson distribution 
barplot(dpois(0:10,3.5),ylab="Probability",xlab="n",ylim=c(0,0.2), names.arg=0:10)
```

# 3.5
```{r}
# Graphing the binomial with n = 20
barplot(dbinom(0:20,20,0.3),ylab="Probability", ylim=c(0,0.2), xlab = 'k', names.arg=0:20)

# This distribution show the probability of getting exactly k successes in n independent Bernoulli trials
```
One important thing that differentiates poisson and binomial distributions:
The binomial distribution is one, whose possible number of outcomes are two, i.e. success or failure. On the other hand, there is no limit of possible outcomes in Poisson distribution
Example of Binomial: coin toss
Example of Poisson: Number of overfloods in a river

# 3.8
```{r}
# Using the fomula on page 43 in ROS
mean1 = 69.1 # Husbands
mean2 = 63.7 # Wifes

sd1 = 2.9 # Husbands
sd2 = 2.7 # Wifes

# Weights
a = 1/2
b = 1/2

print(a*mean1+b*mean2)
print(sqrt((a^2*sd1^2) + (b^2*sd2^2) + (2*a*b*0.3*sd1*sd2))) # WHY IS THIS DIFFERENT
```
# 4.1
Very similar to the example provided the bottom part of page 52 and top of page 53
Give an estimate and standard error of the average effect size.  

500 people got treatment (5$) and 50% of them responded on the survey,
500 people got no treatment and 40% of them responded on the survey.

What is the effect and SE of the treatment. 
Find the standard error of the estimate as:
$$se = \sqrt{(\hat{p}(1-\hat{p})/n} $$
Then use SE the differences $$ \sqrt{se_1^2 + se_2^2} $$

```{r}
n = 1000 ; n1 = 500 ; n2 = 500
p1 = 0.5 ; p2 = 0.4
#Treatment
SE_treat = sqrt(p1*(1-p1)/n)
#Control
SE_control = sqrt(p2*(1-p2)/n)
#The differences.
SE_diff = sqrt(SE_treat^2+SE_control^2)
c(SE_diff = SE_diff, effect_of_treatment = (p1-p2))
```

# 4.2
Using the data on page 53. 

how many people do you need before the se is less than 5 percentage points. 
Well how many people you for a SE to be less than 5 percentage points depends on
the sd if you're working with continuous variables or the $\hat{p}$ if you're working with
proportions. 

```{r}
for (n in 1:300){
  # Proportions
  p_male <- 0.45
  p_female <- 0.57
  
  se_male <- sqrt(p_male*(1-p_male)/n)
  se_female <- sqrt(p_female*(1-p_female)/n)
  
  se_diff <- sqrt(se_male^2+se_female^2)
  if (se_diff < 0.05){
    print(paste(n, "people are required to get a se below 5 percentage points if n is the same for both male and female."))
    break
  }
}

```


hint create a loop through the number of percentage points



