---
title: "Week 1 solution"
author: "Kathrine Schultz-Nielsen"
date: "8/2/2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
pacman::p_load(rstanarm,tidyverse, ggpubr)
```

# 1.2
Predicting percentage vote to incumbent party. A linear relationship is given.
Simulate data with different residual standard deviataions.

```{r}
# y = 46.3 + 3x + error

# rnorm generates a random data set with the given mean and standard deviation
x1=rnorm(1e3,2,2/4) # sampling 1000 datapoint with mu = 2 and sd = 0.5

#Simulate such data with residual sd = 3.9
error=rnorm(1e3,0,3.9) # Ensuring that the residuals (i.e. errors have sd = 3.9)
y1=30+(10*x1)+error 
df_12 <- data_frame(x1 = x1, y1 = y1)
model <- stan_glm(y1~x1, data = df_12, refresh = 0)
#summary(model)
```

```{r}
model1 <- lm(y1~x1, data = df_12)
plot(x = x1, y = y1,)
abline(a = model1$coefficients[1], b = model1$coefficients[2])
```



```{r}
plot(x = x1, y = y1,)
abline(a = model$coefficients[1], b = model$coefficients[2])
```

```{r}
x2=rnorm(1e3,2, 2/4)
#Simulate such data with residual std = 10
error2=rnorm(1e3,0,10)
y2=30+(10*x2)+error2
df_12 <- df_12 %>% 
  mutate(x2 = x2) %>% 
  mutate(y2 = y2)
model2 <- stan_glm(y2~x2, data = df_12, refresh = 0)
#summary(model2)
```


```{r}
plot(x = x2, y = y2,)
abline(a = model2$coefficients[1], b = model2$coefficients[2])
```
#### Better way using a function. 
Create function
```{r}
simulate_data <- function(b,a,sigma,n, max_x){
  b <- c(b, a) #Beta matrix of model parameters
  n <- n #N observations
  x <- runif(n, min = 0, max = max_x) #x
  err <- rnorm(n, mean = 0, sd = sigma) #residual errors
  y <- b[1] + b[2]*x + err
  
  return (tibble(x, y))
}
```

simulate data with sigma = 3.9. 
```{r}
intercept = 30
slope = 10
sim1 = simulate_data(b = intercept, a = slope, sigma =  3.9, n = 2000, max_x = 4)
stan_glm(y ~ x, data = sim1, refresh = 0)
ggplot(sim1, aes(x = x, y = y)) + geom_point() + geom_abline(intercept = intercept, slope = slope, color = "red", size = 1.5)
```

simulate data with sigma = 10.
```{r}
sim2 = simulate_data(b = intercept, a = slope, sigma =  10, n = 2000, max_x = 18)
stan_glm(y ~ x, data = sim2)
ggplot(sim1, aes(x = x, y = y)) + geom_point() + geom_abline(intercept = intercept, slope = slope, color = "red", size = 1.5)
```
## 1.5 Goals of regression
a) Forecasting/classification
- Forecasting stock prices
- Classifying gene expressions
- In cogsci whether some stimuli can predict a behavioral or neural outcome

b) 
- Could be association between global temperature and green house gas emissions

c)
- Prediction of future height in Kids (could be a linear extrapolation), using measures and regression fits to predict future outcomes

d) 
- Does a drug or treatment have an effect compared to a control group


# Chapter 2 

```{r}
remotes::install_github("avehtari/ROS-Examples",subdir = "rpackage")
```

```{r}
# Loading the data
library(rosdata)
```


### Exercise 2.3
```{r}
install.packages("stringr")
pacman::p_load(stringr)
```
```{r}
allnames
```

```{r}
## Creating the data frame
names_f <- filter(allnames, sex == "F") # Filtering all females
names_f <- mutate(names_f, last_letter = str_sub(names_f$name, -1))  # Getting the last letter in their name
names_f <- mutate(names_f, last_letter = as.factor(last_letter)) # Converts to factors
years <- names(allnames)[4:134] # Extracting the year columns
 # group by the last letter
names_f_group <- group_by(names_f, last_letter) # Ensures that future operations are done per group (last letter)

sum_names_f <- summarise_at(names_f_group, years, sum, na.rm = TRUE) # For each group we summarize the number of last letters per year

#See this for overview
#print(names_f)
#print(names_f_group)
#print(sum_names_f)

```

```{r}
#Turning frequencies into percentages
func_1 <- function(x, na.rm = TRUE) (x/sum(x))*100
percentage_names_f<- mutate_at(sum_names_f, years, func_1)
#percentage_names_f
```

```{r}
#Pivot longer so we get 1 row for each year with each letter. 
final_df <- pivot_longer(percentage_names_f, cols = years, names_to = "Year", values_to = "number") 
final_df <- mutate(final_df, Year = as.factor(Year))
final_df <- mutate(final_df, Year = as.numeric(Year))
#final_df
```

```{r}
#Plotting time
final_df %>% 
  ggplot(aes(x = Year, y = number, col = last_letter, group = last_letter)) + geom_line() + 
  labs(y = "Percentage of people born with letter", title = "Overview of % girls born with a specific last letter each year", x = "Years after 1880") +  scale_x_continuous() + theme(axis.text.x = element_text(angle = -45)) 
```

```{r}
#allnames
```

### 2.7

a) An example of measurements that are valid but not reliable:


b) An example of measurements that are reliable but are invalid:
A scale that measure 5 kgs off - it is reliable in the sense that it will measure the same every time, however the measure is invalid as it will be 5 kgs offl





