---
title: "Week 3 solution"
author: "Kathrine Schultz-Nielsen"
date: "22/2/2023"
output:
---

```{r}
# Loading dependencies
pacman::p_load(tidyverse, ggpubr)
```

# Chapter 4

## Warm-up 1: Consider the study of ovulation and political attitudes on pp. 62-63. Why is this study flawed. Discuss with your neighbor :)


## Warm-up 2: What are the authors in ROS 3 suggestions to avoid over-interpretation of noise?


## Optional: If you want more exercises from chapter 4 try 4.1 and 4.2 if you didn't complete them last week. In 4.2 use the proportions given in the example in section "Standard error for a comparison" on pp. 52-53.

# Chapter 5

Before we dig into the exercises from ROS complete the following two small exercises about loops as they will be essential to simulation.

## Exercise 1 Practising loops: What is the difference between a for- and a while loop? Explain in your own words. When would you use for loops and when would you use while loops?

## Exercise 2 Practising loops: Create and run both a simple for and a while loop

```{r}

```

## 5.1
a)
```{r}

```

b) 
```{r}

```
c) 
```{r}

```

## 5.3
a) 
```{r}

```
b) 
```{r}

```
# 5.4
Recall the Central limit theorem on the bottom of page 51.
```{r}

```

# 5.5 
*Hint 1 
```{r}

```


# 5.8
a) 
```{r}

```

b)
*Hint 2

```{r}

```

# 5.9

```{r}

```

# Hints
### Hint 1 
Utilize that 
$$
\rho_{X, Y}=\operatorname{corr}(X, Y)=\frac{\operatorname{cov}(X, Y)}{\sigma_X \sigma_Y}
$$
### Hint 2
If you are stuck figuring out how to find confidence intervals in R have a look here:
https://www.geeksforgeeks.org/how-to-find-confidence-intervals-in-r/
