---
title: "R Notebook"
output: html_notebook
---

```{r}
# importing dependencies
library(tidyverse)
library(rstanarm)
library(bayesplot)
set.seed(0)
```

# 11.5
Residuals and predictions: The folder Pyth contains outcome $y$ and predictors $x_1$, $x_2$ for 40
data points, with a further 20 points with the predictors but no observed outcome. Save the file
to your working directory, then read it into R using read.table()

Remember that you can load the data sets from ROS directly from their github
```{r}
remotes::install_github("avehtari/ROS-Examples",subdir = "rpackage")
library(rosdata)
pyth

```

a) Use R to fit a regression model predicting $y$ from $x_1$ and $x_2$ using the first 40 data points. Use R to summarize inferences and checks the fit of your model. 

Use the function pp_check() to summarize your inference and check your model fit. 

Take a moment to understand what the pp_check() plot is showing you.

b) Display the estimated model graphically as in figure 11.2.

There is code on page 164-165 in ROS that can help you get started

c) Make residual plots for the model. Do the assumptions of the linear model regarding residuals appear to be met?

d) Make predictions for the remaining 20 data points in the file. How confident do you feel
about these predictions?

*hint* use the functions posterior_predict() to create predictive distributions. Use mcmc_intervals() to visualize these.

# 11.9
Use LOO to compare different models fit to the beauty and
teaching evaluations example from Exercise 10.6.
Fit two different models. I have chosen to compare the two following models
$$
eval = \beta_0 + \beta_1 \cdot beauty + \beta_1 \cdot female \\
eval = \beta_0 + \beta_2 \cdot beauty
$$
but feel free to choose models of your own choice

a) Discuss the LOO results for the different models and what this implies, or should imply, for model choice in this example. Hint: look at the looic using the loo() function.

b) Compare predictive errors pointwise. Are there some data points that are having high predictive errors for both fitted models.

*hint* If you have defined the output of the loo() function as a variable called loo2, then you can extract the pointwise looic using loo2$pointwise[,4]. Can you find a way to sort these pointwise looic values in decreasing order and extract their indices?

# 11.3
Consider the following procedure:

Set n = 100 and draw n continuous values x_i uniformly distributed between 0 and 10. Then
simulate data from the model $y_i = a + b \cdot x_i + error_i$, for $i = 1, . . . , n$, with $a = 2$, $b = 3$, and independent errors from a normal distribution.

```{r}
# Your code goes here...
```

Regress $y$ on $x$. Look at the median and mad sd of b. Check to see if the interval formed by
the median ± 2 mad sd includes the true value, b = 3.

```{r}
# Your code goes here...
```

Repeat the above two steps 1000 times.

```{r}
# Your code goes here...
```

a) True or False: The interval should contain the true value approximately 950 times. Explain your answer.


b) True or False:  The interval should contain the true value approximately 950 times IF the error follows a bimodal distribution. Explain your answer.














