---
title: "R Notebook"
output: html_notebook
---


# Matrix calculation by hand. 
exercises are in the folder for today. 

 - Exercise 1, 3 and 6 shown as pictures. 
 
    Solution: see handwritten notes

# Rank and trace 
 - Describe the difference between full rank and short rank
    Determine which matrices from 1 and 6 are full rank
    
 - Determine the trace of the matrices in exercise 1

If you have more time try to complete the exercises 1, 3 and 6 using matrix operations in R
At least explore the cells below

# Intro to matrix operations in R.
```{r}
pacman::p_load(tidyverse)
```

*Multiplication*
```{r}
matrix1 <- matrix(seq(1,10, by = 1), ncol = 2)
matrix2 <- matrix(seq(1,10, by = 1), nrow = 2)
# 5x2 premultiplied by 2x5 = 5x5
matrix1 %*% matrix2
# 2x5 premultiplied by 5x2 = 2x2 
matrix2 %*% matrix1
```

*Transposing*
```{r}
matrix1
# We can transpose using t()
t(matrix1)
```
The first column is now the first row and so forth. 


*Inverse*
```{r}
# solve() is used to find the inverse of a matrix
solve(matrix1)
```
This gives us an error message that saying that we don't have a square matrix and therefore cannot find the inverse. 

A matrix multiplied by its own transposed will always give us a square matrix. 
```{r}
t(matrix1) %*% matrix1
```

```{r}
matrix1 %*% t(matrix1)
```
But as we can see the order still maters. If we take t(X) pre-multiplied by X we will get a square matrix with the size of the original X's number of columns. If the order of multiplication is inversed we will of course get the new matrix with dimensions equal 
to the original X's number of rows.

We find the inverse of $(t(X)X)^{-1}$
```{r}
solve(t(matrix1) %*% matrix1)
```
Another issue that can arise with inversions of matrices. 
```{r}
matrix3 <- matrix( c(1,2,3,6) ,ncol = 2, nrow = 2)
matrix3
```
We see that this matrix is a square matrix so we should be able to find the inverse,
or should we?? Let's try it. 

```{r}
solve(matrix3)
```
It returns the message that is singular. What does that mean? How does it relate to the 
our knowledge of the determinant? 

*Determinant*
We can find the determinant of matrix using the following det() function. 
```{r}
det(matrix3)
```
The problem here is that the matrix does not have full rank

